---
title: "Untitled"
output: html_document
date: "2023-10-24"
---

They have provided you with a single dataset to use. A summary and preview is provided below.
The data is publicly available here: (https://data.lacity.org/Public-Safety/Crime-Data-from-2020-to-Present/2nrs-mtv8)
👮‍♀️ crimes.csv

```{r, results=FALSE, message=FALSE, warning=FALSE, include=FALSE}
library(tidyverse)
crimes <- suppressWarnings(read_csv("C:/Users/nici_/Downloads/crimes.csv"))
head(crimes)
```

```{r, results=FALSE, message=FALSE, warning=FALSE, include=TRUE}
library(tidyverse) 
library(dplyr)

#Filter all with no missings
crimes2 <- crimes |> rename("descent" = "Vict Descent")
crimes2 <- crimes2 |> rename("sex" = "Vict Sex")
crimes2 <- crimes2 |> rename("age" = "Vict Age")
crimes2 <- crimes2 |> rename("area" = "AREA NAME")
crimes2 <- crimes2 |> rename("premis" = "Premis Desc")

#crime category
crimes2 <- crimes2 |> rename("crm" = "Crm Cd")

crimes2 <- crimes2 |> filter(!is.na(descent))
crimes2 <- crimes2 |> filter(!is.na(sex))

crimes2 <- crimes2 |> 
 filter(descent %in% c("A", "B", "C", "D", "F", "G", "H", "I", "J", "K", "L", "O", "P", "S", "U", "V", "W", "X", "Z"))

crimes2 <- crimes2 |> 
 filter(sex %in% c("F", "M"))

crimes2 <- crimes2 |> 
 filter(age >= 0)
```

## Grouped Crime Categories

- 1 Homicide: Code 110, 113 
- 2 Rape: Code 121, 122, 815, 820, 821 
- 3 Robbery: Code 210, 220
- 4 Aggressive assaults: Code 230, 231, 235, 236, 250, 251, 761, 926
- 5 Simple assault: Code 435, 436, 437, 622, 623, 624, 625, 626, 627, 647, 763, 928, 930 
- 6 Bulgary: Code 310, 320

```{r, results=FALSE, message=FALSE, warning=FALSE, include=TRUE}
library(mgcv)

crimes2$sex <- as.factor(crimes2$sex)
crimes2$descent <- as.factor(crimes2$descent)
crimes2$area <- as.factor(crimes2$area)
crimes2$premis <- as.factor(crimes2$premis)


#group for crime
crimes2 <- crimes2 |> 
  dplyr::mutate(serious = case_when(crm %in% c("110", "113") ~ "1", #homicide
                                    crm %in% c("121", "122", "815", "820", "821") ~ "2", #rape
                                    crm %in% c("210", "220") ~ "3", #robbery
                                    crm %in% c("230", "231", "235", "236", "250", "251", "761", "926") ~ "4", #aggassaults
                                    crm %in% c("435", "436", "437", "622", "623", "624", "625", "626", "627", "647", "763", "928", "930") ~ "5", #sassault
                                    crm %in% c("310", "320") ~ "6", TRUE ~ "0")) #bulgary

crimes2 <- crimes2 |> filter(serious != "0")
crimes2$serious <- as.factor(crimes2$serious)


#class Victims in age groups
crimes2 <- crimes2 |> mutate(age20 = case_when(age>=0 & age<=18 ~ "minor", TRUE ~ "adult"))

crimes2$age20 <- as.factor(crimes2$age20)
```

The victims descent will be grouped into broader descent categories:

Asian:
C - Chinese
J - Japanese
K - Korean
L - Laotian
V - Vietnamese
Z - Asian Indian
A - Other Asian

Black:
B - Black

Native American/Alaskan Native:
I - American Indian/Alaskan Native

Pacific Islander:
G - Guamanian
P - Pacific Islander
S - Samoan
U - Hawaiian

Hispanic/Latin/Mexican:
H - Hispanic/Latin/Mexican

White:
W - White

Other:
D - Cambodian
F - Filipino
O - Other
X - Unknown

This grouping should make it easier to analyze and understand the data based on these broader categories.

```{r, results=FALSE, message=FALSE, warning=FALSE, include=TRUE}
#group victims origin into broader groups
crimes2 <- crimes2 |> 
 mutate(descent_g = case_when(descent %in% c("A", "C","J", "K", "L", "V", "Z") ~ "Asian", 
descent == "B" ~ "Black", 
descent == "I" ~ "Native American",
descent %in% c("G", "P", "S", "U") ~ "Pacific Islander", 
descent == "H" ~ "Hispanic", 
descent == "W" ~ "White", 
TRUE ~ "Other"))

crimes2$descent_g <- as.factor(crimes2$descent_g)
```

```{r, results=FALSE, message=FALSE, warning=FALSE, include=TRUE}
#turn dates into season
library(hydroTSM)

crimes2$seasons <- time2season(as.Date(crimes2$`Date Rptd`), # Convert dates to seasons
                          out.fmt = "seasons")

crimes2 <- crimes2 |> filter(!is.na(seasons))

```

```{r, results=TRUE, message=FALSE, warning=FALSE, include=TRUE}
library(caret)
library(sjPlot)
library(sjmisc)
library(sjlabelled)


#partition data frame into training and testing sets
#set labels
crimes2$serious <- set_labels(crimes2$serious, labels = c("Homicide", "Rape", "Robbery", "Aggressive Assaults", "Simple Assault", "Bulgary"))
train_indices <- createDataPartition(crimes2$serious, times=1, p=.8, list=FALSE)

#create training set
df_train <- crimes2[train_indices , ]

#create testing set
df_test  <- crimes2[-train_indices, ]


#multinominal logistic Regression for category of the commited crime
model <- nnet::multinom(serious ~ age + sex:descent_g + seasons + area, data=df_train, MaxNWts=100000)

model_list <- as.list(model)
tab_model(model_list, p.style = "numeric_stars")
```

```{r, results=TRUE, message=FALSE, warning=FALSE, include=TRUE}
#plot log odds
library(ggthemes)
library(nnet)
library(ggeffects)

gen <- ggpredict(model, terms = "seasons")
plot(gen, facet = TRUE, colors = "metro")

min <- ggpredict(model, terms = "age")
plot(min, facet = TRUE, colors = "metro")

des <- ggpredict(model, terms = "sex:descent_g")
plot(des, facet = TRUE, colors = "metro") 
```

```{r, results=TRUE, message=FALSE, warning=FALSE, include=TRUE}
#prediction
predict <- predict(model, new.data = df_test)
predict_df <- as.data.frame(predict)
predict_df <- set_labels(predict_df, labels = c("Homicide", "Rape", "Robbery", "Aggressive Assaults", "Simple Assault", "Bulgary"))

#summary
view_df(predict_df, show.frq = T, show.prc = T)
view_df(df_test, show.frq = T, show.prc = T)

```

# Geo-Cluster
```{r, results=TRUE, message=FALSE, warning=FALSE, include=TRUE}
library(maps)
library(tidyverse)
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)

crimes_map <- crimes2 |> filter(!is.na(LON) & LON !=0)
crimes_map <- crimes_map |> filter(!is.na(LAT) & LAT !=0)

crimes_map <- crimes2 |> filter(!is.na(LON) & LON !=0)
crimes_map <- crimes_map |> filter(!is.na(LAT) & LAT !=0)

#filter Simple assault

homicide <- crimes_map |> filter(serious == 1)
# Get long and lat from data.frame for map

lon <- c(homicide$LON)
lat <- c(homicide$LAT)
df <- as.data.frame(cbind(lon,lat))

sites <- data.frame(longitude = -118.243683, latitude = 34.052235)

world <- ne_countries(scale = "medium", returnclass = "sf")
class(world)


#Map Homicide
base_map <- ggplot(data= world, width = 50, height = 36) +
    geom_sf() +
    geom_point(data = df, aes(x = lon, y = lat), size = 2, alpha = .15,  
        shape = 23, fill = "darkred") +
    coord_sf(xlim=c(-118.0, -118.7), ylim = c(33.7, 34.35),, expand = TRUE) + theme_classic()

```
